Ligeti, Stockhausen and Koenig:\\Aesthetics of Electronic Music

### 2015-09-23 manuscript.tex
Add 'Comparison' subsection to 'concepts'
Koenig ideas

>>>>

``The Construction of Sound'' by Koenig (\cite*{koenig}) deals mainly with serial music and in different ways that serial decisions can be incorporated into electronic music.
The current work explores aesthetics in electronic music.
Thus, in this section I try to extract electronic music characteristics from the serial music context in which they are presented in the paper.
As before, I quote from the original paper whenever possible and add my own clarifications only when necessary.

\subsubsection{Sounds of limited vs. unlimited lengths}
\label{subs:koenig:lengths}

\begin{MyShadequote}
  (Intrumental music characterized by) homogeneous, continuous sounds of limited lengths.
  One might say that all classic categories of melodic formation, counterpoint, harmony, form and instrumentation have their origins in this manner of production.
  (On the other hand), the length of electronically produced sound is theoretically unlimited...
  The classic categories thus do not apply to electronic music;
  at any rate they do not have their origins in the new material and the way in which it is produced (p. 1-2).
\end{MyShadequote}

\subsubsection{Non-stationary sounds}
\label{subs:koenig:nonstationary}

Koenig classify timbre into ``classes''.
Both instrumental and electronic music share three classes of continuous sounds:
Harmonic sound, which consists a fundamental; Noise sound, without fundamental; and impulses, which have no timbre nor fundamental.
According to Koenig ``(for the three classes above) we have one timbre for each attack'' (p. 17).

Electronic music enhances the timbre possibilities with another class:

\begin{MyShadequote}
 Electronic music also knows a family of non-stationary sounds with a duration, but no constant timbre...
 Since their ``timbre'' is not constant, it can be described but not accurately imagined.
 The movement of the timbre rather follows the compositional process of the micro-time range.
 In this ways structures are composed instead of timbre (p. 18).
\end{MyShadequote}

\subsubsection{Sound as a formal event}
\label{subs:koenig:event}

\begin{MyShadequote}
  The electronically produced sound does not have to be homogeneous and continuous.
  One complete sound can consist of several individual components whose determinants may change in the next sound.
  The sound is a closed unit and becomes a means of formation;
  each sound characteristic becomes a formal event.
  Instrumental music composes ``with'' sounds, electronic music composes sounds (p. 2).
\end{MyShadequote}

\subsubsection{Representation of pitch ranges}
\label{subs:koenig:pitch_ranges}

This paper deals mainly with serializm, and how it is applied to electronic music.
The following idea came from this way of thinking, in serias, but I believe that it might by understood and analyzed even without considering serial decisions.
Here, Koenig describe different ways to represent a pitch range.

\begin{MyShadequote}
  \begin{description}
    \item[Pointillistically] - single pitches are used in such a way that the limits of the range are frequently touched.

    \item[In planes] - the range is filled as continuously as possible...
    The range as a whole is clear, since limits and interior are uniformly occupied.

    \item[Articulated in planes] - the entire plane of the range is divided up into changing sections which in their turn can be either pointillist or planes:
    pointillist as ``chords'', clusters of various sizes;
    planes in the form of smaller planes within the larger one.

    \item[Transition] - it is a simple matter to have a regular transition from planes to points if the plane is brought about by a rapid sequence of points by:
    (a) slowing down the process,
    or (b) reducing the number of layers step by step (p. 11).
  \end{description}

\end{MyShadequote}

\subsection{Comparison}

TODO regarding the first concept by Stockhausen:
Overall, I can not find a common ground between this compository element to those of Ligeti that were mentioned before.
Their attitudes regarding unified perception of time are, most probably, contradicting.
